On Cybersecurity\\Science and Engineering\footnote{Confidential, restricted to the authors. Property of V-Research S.r.l.}

### 2020-02-19 v-research_experiment-alpha.tex
extend with exploitation flow

\subsubsection{Process: Incidents and Accidents}\label{sec:incidentaccident}
In our informal definition depicted in Figure~\ref{fig:safety-security}, a
Vulnerability is seen as a process through which an agent (i.e. Mankind)
generates an Incident, i.e.
Agent$\rightarrow$Vulnerability$\rightarrow$Exploit$\rightarrow$Attack$\rightarrow$Incident.
We describe processes, in our current formalization, in terms of the K-Modal
Structure as in Definition~\ref{def:modalinterpretation} (i.e. in terms of
causality principle).

@@

\begin{example}
given in natural language in the description of the 
CWE-116 which we can analyze as follows:
\begin{itemize}
	\item ``the software prepares a structured message''
	\item [the message is intended] ``for communication with another component''
	\item ``but''
	\item ``encoding or escaping of the data is either missing or done incorrectly''
	\item ``[\ldots] result, the intended structure of the message is not preserved''
\end{itemize}
\end{example}

>>>>

\subsubsection{Process: Incidents and Accidents}\label{sec:incidentaccident}
In our informal definition depicted in Figure~\ref{fig:safety-security}, an
Incident is generated through a process caused by a Vulnerability (which,
in turn, is caused by a Weakness). Symmetrically, Nature has a process from Defect
to Accidents. We start by analyzing the process generated by Mankind, which is
divided into the following phases (explained in the remainder of this sub-Section):
%Agent$\rightarrow$Vulnerability$\rightarrow$Exploit$\rightarrow$Attack$\rightarrow$Incident.
\begin{enumerate}[noitemsep]
	\item An agent (Mankind or Nature) defines a system, e.g. by following an engineering process for creating a new system, or for redesigning or improving an existing system (e.g. a component or a collection of components) 
	\item The dogmatic definition of a system contains a Weakness (i.e. the agent, Mankind or Nature, introduces an unintended behavior), for example as an un-intended behavior or as an architectural fallacy. The Weakness can be applied to the abstract definition of the system.
	\item The Weakness can be defined as a malicious process for a system, called Vulnerability (or Defect for Nature)
	\item The process expressed by the Vulnerability can be made operational for a specific implementation of the system (i.e. implemented by Mankind or realized by Nature) 
\end{enumerate}
Those additional details are depicted in Figure~\ref{fig:safety-security_2}, an update
of Figure~\ref{fig:safety-security}.

\begin{figure}[t]
	\centering
	\includegraphics[width=\textwidth]{safety-security_2.pdf}
	\caption{Overview security and safety keywords}
	\label{fig:safety-security_2}
\end{figure}

\Paragraph{Exploitation Flow} An \emph{Exploit} is an \emph{implementation} of
a Vulnerability, where a Vulnerability is one of the many ways to transform a
Weakness into a malicious process that can be applied\footnote{Meaning that the
malicious process, i.e.  any structured procedure such as a protocol, can be
somehow made operational, e.g. implemented} to one or many systems. So, the act
of an agent (Mankind) of ``exploiting a Vulnerability'' is the process of
making the Vulnerability operational, by implementing the Vulnerability w.r.t.
a specific target system.  Whenever the agent is Nature, the implementation is
more broadly considered a realization\footnote{``The state of being
realized''\autocite{Merriam2020realization}}

The Common Attack Pattern Enumeration and
Classification\autocite{MITRE2020CAPEC} (CAPEC) online service provides a
``community resource'' which aims at ``identifying and understanding attacks''.
The MITRE writes on the CAPEC homepage that ``Understanding how the adversary
operates is essential to effective cyber security. CAPEC helps by providing a
comprehensive dictionary of known patterns of attack employed by adversaries to
exploit known weaknesses in cyber-enabled capabilities. It can be used by
analysts, developers, testers, and educators to advance community understanding
and enhance defenses.''. We can extrapolate two major concepts: 
\begin{enumerate}
	\item while understanding the exploitation process is the aim of the
		CWE and CVE initiatives, understanding the hacking process is
		the aim of the CAPEC initiative.
	\item CAPEC provides a \emph{dictionary} of \emph{known} attack patterns.
		As stated largely in Section~\ref{sec:}, this dictionary contains
		empirical evidences and should not be used to induce a cybersecurity 
		theory; but it can be used to validate a theory.
\end{enumerate}

The process continues as follows.
\begin{enumerate}[noitemsep]
\setcounter{enumi}{4}
	\item an agent (Mankind\footnote{For the sake of readability we focus only on Mankind as an agent}) the exploitation flow can be used to hack a system. We use the term hack (as hacking) to stress that nothing prevents this process from being honest, however, for the sake of simplicity, we focus on Attacks which are malicious by definition.
	\item An Attack is applied to a system with malicious intent by an agent
	\item The application of the attack results in an Incident which impacts, i.e. poses a Threat to, an Asset
\end{enumerate}

\Paragraph{Hacking Flow}

\begin{example}
given in natural language in the description of the 
CWE-116 which we can analyze as follows:
\begin{itemize}
	\item ``the software prepares a structured message''
	\item the message is intended ``for communication with another component''
	\item ``but''
	\item ``encoding or escaping of the data is either missing or done incorrectly''
	\item ``[\ldots] result, the intended structure of the message is not preserved''
\end{itemize}
\end{example}

In this Vulnerability is seen as a process through which an agent (i.e. Mankind)
generates an Incident, i.e.

We describe processes, in our current formalization, in terms of the K-Modal
Structure as in Definition~\ref{def:modalinterpretation} (i.e. in terms of
causality principle).

@@
